---
title: "Generate imputed data to use when creating IPW for genteic testing survey"
output: 
  html_document:
    toc: true
    keep_md: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r,message=FALSE}
library(tidyverse)
library(survey)
library(plyr)
library(dplyr)
library(factoextra)
library(labelled)
library(tableone)
library(memisc)
library(sjlabelled)
library(mi)
library(boot)
```

# Set up working directory and packages

```{r}
 
base_path <- "~/Dropbox (Partners HealthCare)/SOL_misc_genetics/survey_gen_test_utilization/20210917_data_code/"

```


# Overview
This code is to impute variables which will then be used in generating IPW for the genetic testing survey. 
Only a subset of HCHS/SOL individuals who participated in visit 1, participated 
in the genetic testing survey. We cannot use visit 2 weights, because some of the
people who participated in the genetic testing survey, did not participate in visit 2.
Therefore, we estimate the probability of participating in the genetic testing survey 
as follows: 
1. Estimate the probability of participating in the annual follow up phone call (AFU)
of 2019-2020, given that as individual participated in HCHS/SOL visit 1. 
2. Estimate the probability of responding to the genetic testing survey, given
that an individual participated in the AFU. 

There is sporadic missingness, so we first impute missing data, based on visit 1 only. For individuals who have more updated covariates from the AFU or from visit 2, we will use these more updated data when estimating the probability of responding to the genetic testing survey, and for association analysis.



# Creating dataset for imputation
First, we need to identify people who participated in the AFU.
```{r}
# read the genetic testing survey data: 
survey_gen <- read.csv(paste0(base_path, "/Preprocessed_data/solfu_gte_20200730.csv"))

# remove rows with data from afu_year =2 
survey_gen <- survey_gen[-which(survey_gen$AFU_YEAR==2),]
head(survey_gen)

```
In the survey data, there are `r nrow(survey_gen)` rows, corresponding to  `r length(unique(survey_gen$ID))` individuals. 


```{r}
# read covariates: 
covariates <- read.csv(paste0(base_path,"Preprocessed_data/solfu_covariates_20210917.csv"))
head(covariates)

# create a variable AFU_par that is equal to 1 if a person participated in the AFU:
covariates$AFU_par <- ifelse(is.element(covariates$ID, survey_gen$ID), 1, 0 )
```

Of `r nrow(covariates)` HCHS/SOL participants, `r sum(covariates$AFU_par)` participated in the AFU, and `r sum(covariates$AFU_par ==0)` did not. 


# Prepare variable "physician_visits" 
In the dataset, there are two forms of "HCE5" (how many times saw a physician in the last 12 months): "HCEA5" and "HCEB5" that record different participants. We will combine two of them and refill the missing in HCEA5 with the value of HCEB5 with the same ID.
Afterwards, we re-level the HCEA5 and sort them into 3 groups: 
(1)If HCEA5==0, then it is "No".
(2)If HCEA5 ==1 or 2, then it is "One or Two times". 
(3)If HCEA5>=3, it is "At least three times".
```{r}
covariates$HCEA5[covariates$HCEA5==""] <- NA
covariates$HCEB5[covariates$HCEB5==""] <- NA

covariates$HCEA5 <- ifelse(is.na(covariates$HCEA5)==TRUE,covariates$HCEB5,covariates$HCEA5)
names(covariates)[names(covariates)=="HCEB5"] <- "HCE5"

covariates$HCEA5 <- as.numeric(covariates$HCEA5)
covariates$physician_visits <- ifelse(covariates$HCEA5>=3,2,covariates$HCEA5)
covariates$physician_visits <- ifelse(covariates$HCEA5==2,1,covariates$physician_visits)

# sum(is.na(covariates$HCE5))

```



# Categorize variables and create factors
In this section, we factorize all variables for imputation and logistic regression and define levels for all variables to be used. We re-defined levels of two variables: "Employment_status" and "age". 

```{r}
covariates_1 <- covariates

covariates_1 <- covariates_1 %>%
  mutate(
     SEX=factor(GENDER,levels = c("F","M"),labels=c("Female","Male")),
    Education=factor(EDUCATION_C3,levels = c("1","2","3"),labels=c("<12","12",">12")),
    CENTER=factor(CENTER,levels = c("B","C","M","S"),labels=c("Bronx","Chicago", "Miami", "San Diego")),
    Income_level=factor(INCOME_C5,levels = c("1","2","3","4", "5"),labels = c("Less than $10,000","$10,001-$20,000","$20,001-$40,000","$40,001-$75,000","More than $75,000")),
    Current_Health_insurance=factor(N_HC,levels = c("0","1"),labels = c("No","Yes")),
    Background=factor(BKGRD1_C7,levels = c("0","1","2","3","4", "5","6"),labels = c("Domician","Central American","Cuban","Mexican","Puerto Rican","South American","More than one/Other heritage")),
    Language_pref = factor(LANG_PREF, levels = c("1", "2"), labels = c("Spanish", "English")),
    Marital_status=factor(MARITAL_STATUS,levels=c("1","2","3"),labels = c(
      "Single","Married or living with a partner","Separated,divorced,or widow(er)")),
    Employment_status=factor(EMPLOYED, levels=c("1","2","3","4"),labels = c("Retired/not currently employed","Not retired and not currently employed","Employed part-time(<=35 hours/week)","Employed full-time(>35 hours/week)")),
    US_BORN=factor(US_BORN, levels=c("0","1"),labels = c("No","Yes")),
    Physician_Visit=factor(physician_visits,levels = c("0","1","2"), labels = c("No","One or two times","At least three times"))
  )



covariates_1$Employment_status <- relabel(
  covariates_1$Employment_status,
  "Retired/not currently employed" ="Retired/not currently employed",
  "Not retired and not currently employed"="Retired/not currently employed"
)


covariates_1$age_cat <- case_when(covariates_1$AGE <= 25 ~ '<25',
                              between(covariates_1$AGE, 26,40) ~ '26-40',
                              between(covariates_1$AGE, 41,60) ~ '41-60',
                            covariates_1$AGE >= 61 ~ '61<')

covariates_1$age_cat <- as.factor(covariates_1$age_cat)
saveRDS(covariates_1,paste0(base_path,"Processed_data/20210917_organized_visit1_covariates.RData"))
```

The levels of "Employment_status" is `r get_labels(covariates_1$Employment_status)`. The levels of "age_cat" are `r get_labels(covariates_1$age_cat)`



# Performing Imputation
First, we extract the variables to be applied in the imputation and then create a data matrix. We repeat the process 5 times. 
```{r}
set.seed(800)
predictors <- c("age_cat",
                             "SEX",
                             "Education",
                             "CENTER",
                             "Income_level",
                             "Current_Health_insurance",
                             "Physician_Visit",
                             "Background", 
                             "Language_pref", 
                             "Marital_status",
                             "Employment_status", 
                             "US_BORN",
                             "STRAT")
  
imputedat <- covariates_1[,c("ID", predictors)]
cl <- parallel::makeCluster(5, setup_strategy = "sequential")
## Revert to 'sequential' setup of PSOCK cluster in RStudio Console on macOS and R 4.0.0
if (Sys.getenv("RSTUDIO") == "1" && !nzchar(Sys.getenv("RSTUDIO_TERM")) &&
    Sys.info()["sysname"] == "Darwin" && getRversion() >= "4.0.0") {
  parallel:::setDefaultClusterOptions(setup_strategy = "sequential")
}

mdf <- missing_data.frame(imputedat)
mdf1 <- mi(mdf)
imputed_data <- as.data.frame(complete(mdf1,m=5))
saveRDS(imputed_data,paste0(base_path,"Processed_data/20210917_visit1_imputed_data.RData"))

```




