---
title: "Generate imputed data to use when creating IPW for genetic testing survey (part 2)"
output: 
  html_document:
    toc: true
    keep_md: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r,message=FALSE}
library(tidyverse)
library(survey)
library(plyr)
library(dplyr)
library(factoextra)
library(labelled)
library(tableone)
library(memisc)
library(sjlabelled)
library(mi)
library(boot)
```

# Set up working directory and packages

```{r}
 
 base_path <- "~/Dropbox (Partners HealthCare)/SOL_misc_genetics/survey_gen_test_utilization/20210917_data_code/"

```


# Overview
This code is to impute variables which will then be used in generating IPW for the genetic testing survey. 
Only a subset of HCHS/SOL individuals who participated in visit 1, participated 
in the genetic testing survey. We cannot use visit 2 weights, because some of the
people who participated in the genetic testing survey, did not participate in visit 2.
Therefore, we estimate the probability of participating in the genetic testing survey 
as follows: 
1. Estimate the probability of participating in the annual follow up phone call (AFU)
of 2019-2020, given that individual participated in HCHS/SOL visit 1. 
2. Estimate the probability of responding to the genetic testing survey, given
that an individual participated in the AFU. 

There is sporadic missingness. We first imputed missing data, based on visit 1 only, to use when computing probabilities of participation in the AFU. In this report we are imputing missing data after updating various covariates based on data from visit 2 (or, in the case of marital status, from the 2nd AFU), for use when computing probabilities of participating in the genetic testing survey.



# Creating dataset for imputation
First, we need to identify people who participated in the AFU.
```{r}
# read the genetic testing survey data: 
survey_gen <- read.csv(paste0(base_path, "/Preprocessed_data/solfu_gte_20200730.csv"))

# remove rows with data from afu_year =2 
survey_gen <- survey_gen[-which(survey_gen$AFU_YEAR==2),]
head(survey_gen)

```
In the survey data, there are `r nrow(survey_gen)` rows, corresponding to  `r length(unique(survey_gen$ID))` individuals. 


```{r}
# read covariates: 
covariates <- read.csv(paste0(base_path,"Preprocessed_data/solfu_covariates_20210917.csv"))
head(covariates)

# create a variable AFU_par that is equal to 1 if a person participated in the AFU:
covariates$AFU_par <- ifelse(is.element(covariates$ID, survey_gen$ID), 1, 0 )
```

Of `r nrow(covariates)` HCHS/SOL participants, `r sum(covariates$AFU_par)` participated in the AFU, and `r sum(covariates$AFU_par ==0)` did not. 


## Prepare variable "physician_visits" 
In the dataset, there are two forms of "HCE5" (how many times saw a physician in the last 12 month): "HCEA5" and "HCEB5" that records different participants. We will combine two of them and refill the missing in HCEA5 with the value of HCEB5 with the same ID.
Afterwards, we re-level the HCEA5 and sort them into 3 groups: 
(1)If HCEA5==0, then it is "No".
(2)If HCEA5 ==1 or 2, then it is "One or Two times". 
(3)If HCEA5>=3, it is "At least three times".
```{r}
covariates$HCEA5[covariates$HCEA5==""] <- NA
covariates$HCEB5[covariates$HCEB5==""] <- NA

covariates$HCEA5 <- ifelse(is.na(covariates$HCEA5)==TRUE,covariates$HCEB5,covariates$HCEA5)
names(covariates)[names(covariates)=="HCEB5"] <- "HCE5"

covariates$HCEA5 <- as.numeric(covariates$HCEA5)
covariates$physician_visits <- ifelse(covariates$HCEA5>=3,2,covariates$HCEA5)
covariates$physician_visits <- ifelse(covariates$HCEA5==2,1,covariates$physician_visits)

# sum(is.na(covariates$HCE5))

```

## Update marital status variable
Marital status was available from the baseline visit and year 2 AFU. So when possible, we use the information from year 2 AFU, because it is more recent:

```{r}
survey_marital <- read.csv(paste0(base_path, "/Preprocessed_data/solfu_gte_20200730.csv"))
# keep rows with data from afu_year =2 
survey_marital <- survey_marital[which(survey_marital$AFU_YEAR==2),]
# re-level these to be the same as the collapsed marital_status variable:
# separated, divorced, or widower should become 3
survey_marital$OTEA52[which(is.element(survey_marital$OTEA52, c("1","2","3")))] <- 3
# single should become 1
survey_marital$OTEA52[which(survey_marital$OTEA52 == "4")] <- 1 
# married or living with a partner should become 2
survey_marital$OTEA52[which(is.element(survey_marital$OTEA52, c("0","5")) )] <- 2 
survey_marital$OTEA52[which(survey_marital$OTEA52 == "S")] <- NA
survey_marital$OTEA52[which(survey_marital$OTEA52 == "")] <- NA

covariates <-merge(survey_marital[,c("ID", "OTEA52")],covariates, by="ID", all=TRUE)
inds_updated <- which(!is.na(covariates$OTEA52))
covariates$MARITAL_STATUS[inds_updated] <- covariates$OTEA52[inds_updated]

covariates$OTEA52 <- NULL
 
```


## Update variables to have visit 2 data when available 
Additional variables are available from visit 2. So when possible, we use the information from visit 2, because it is more recent:

```{r}
# Health insurance:
inds_update <- which(!is.na(covariates$N_HC_V2))
covariates$N_HC[inds_update] <- covariates$N_HC_V2[inds_update]

# Education:
inds_update <- which(!is.na(covariates$EDUCATION_C3_V2))
covariates$EDUCATION_C3[inds_update] <- covariates$EDUCATION_C3_V2[inds_update]

# Household income:
inds_update <- which(!is.na(covariates$INCOME_C5_V2))
covariates$INCOME_C5[inds_update] <- covariates$INCOME_C5_V2[inds_update]

# Employment status:
inds_update <- which(!is.na(covariates$EMPLOYED_V2))
covariates$EMPLOYED[inds_update] <- covariates$EMPLOYED_V2[inds_update]

# Age: use age at the survey when possible
# first, reduce the data to that from the AFU

covariates_AFU <- covariates[which(covariates$AFU_par == 1),]
# in the AFU...
inds_na_ageafu <- which(is.na(survey_gen$ageafu)) 
length(inds_na_ageafu)
survey_gen$ageafu[inds_na_ageafu] <- survey_gen$AGE[inds_na_ageafu] + survey_gen$AFU_YEAR[inds_na_ageafu]

# there are still some missing values... 
inds_na_ageafu <- which(is.na(survey_gen$ageafu))
length(inds_na_ageafu)
survey_gen$ageafu[inds_na_ageafu] <- survey_gen$AGE[inds_na_ageafu] + median(survey_gen$AFU_YEAR, na.rm = TRUE)

covariates_AFU <- merge(covariates_AFU, survey_gen[-which(duplicated(survey_gen$ID)),c("ID", "ageafu")])

```


# Categorize variables and create factors
In this section, we factorize all variables for imputation and logistic regression and define levels for all variables to be used. We re-defined levels of three variables: "Race","Employment_status" and "age". 

```{r}
covariates_1 <- covariates_AFU

covariates_1 <- covariates_1 %>%
  mutate(
     SEX=factor(GENDER,levels = c("F","M"),labels=c("Female","Male")),
    Education=factor(EDUCATION_C3,levels = c("1","2","3"),labels=c("<12","12",">12")),
    CENTER=factor(CENTER,levels = c("B","C","M","S"),labels=c("Bronx","Chicago", "Miami", "San Diego")),
    Income_level=factor(INCOME_C5,levels = c("1","2","3","4", "5"),labels = c("Less than $10,000","$10,001-$20,000","$20,001-$40,000","$40,001-$75,000","More than $75,000")),
    Current_Health_insurance=factor(N_HC,levels = c("0","1"),labels = c("No","Yes")),
    Background=factor(BKGRD1_C7,levels = c("0","1","2","3","4", "5","6"),labels = c("Domician","Central American","Cuban","Mexican","Puerto Rican","South American","More than one/Other heritage")),
    Language_pref = factor(LANG_PREF, levels = c("1", "2"), labels = c("Spanish", "English")),
    Marital_status=factor(MARITAL_STATUS,levels=c("1","2","3"),labels = c(
      "Single","Married or living with a partner","Separated,divorced,or widow(er)")),
    Employment_status=factor(EMPLOYED, levels=c("1","2","3","4"),labels = c("Retired/not currently employed","Not retired and not currently employed","Employed part-time(<=35 hours/week)","Employed full-time(>35 hours/week)")),
    US_BORN=factor(US_BORN, levels=c("0","1"),labels = c("No","Yes")),
    Physician_Visit=factor(physician_visits,levels = c("0","1","2"), labels = c("No","One or two times","At least three times"))
  )


covariates_1$Employment_status <- relabel(
  covariates_1$Employment_status,
  "Retired/not currently employed" ="Retired/not currently employed",
  "Not retired and not currently employed"="Retired/not currently employed"
)


covariates_1$age_cat <- case_when(covariates_1$ageafu <= 40 ~ '<40',
                              between(covariates_1$ageafu, 41,60) ~ '41-60',
                            covariates_1$ageafu >= 61 ~ '61<')

covariates_1$age_cat <- as.factor(covariates_1$age_cat)
saveRDS(covariates_1,paste0(base_path,"Processed_data/20210918_organized_covariates_updated_from_vis2.RData"))
```



# Performing Imputation
First, we extract the variables to be applied in the imputation and then create a data matrix. We repeat the process 5 times. 
```{r}
set.seed(800)
predictors <- c("age_cat",
                             "SEX",
                             "Education",
                             "CENTER",
                             "Income_level",
                             "Current_Health_insurance",
                             "Physician_Visit",
                             "Background", 
                             "Language_pref", 
                             "Marital_status",
                             "Employment_status", 
                             "US_BORN",
                             "STRAT")
  
imputedat <- covariates_1[,c("ID", predictors)]
cl <- parallel::makeCluster(5, setup_strategy = "sequential")
## Revert to 'sequential' setup of PSOCK cluster in RStudio Console on macOS and R 4.0.0
if (Sys.getenv("RSTUDIO") == "1" && !nzchar(Sys.getenv("RSTUDIO_TERM")) &&
    Sys.info()["sysname"] == "Darwin" && getRversion() >= "4.0.0") {
  parallel:::setDefaultClusterOptions(setup_strategy = "sequential")
}

mdf <- missing_data.frame(imputedat)
mdf1 <- mi(mdf)
imputed_data <- as.data.frame(complete(mdf1,m=5))
saveRDS(imputed_data,paste0(base_path,"Processed_data/20210918_imputed_data_updated_from_vis2.RData"))

```

# Package versions
```{r}
sessionInfo()
```



