---
title: "Logistic regression of genetic testing awareness"
output: 
  html_document:
    toc: true
    keep_md: true
---


```{r}
library(tidyverse)
library(survey)
library(plyr)
library(dplyr)
library(factoextra)
library(labelled)
library(memisc)
library(jtools)
library(Publish)
library(gtsummary)
library(flextable)
```

# Setting up working directory
```{r}
base_path <- "~/Dropbox (Partners HealthCare)/SOL_misc_genetics/survey_gen_test_utilization/20221003_data_code/"
data_file <- paste0(base_path, "Processed_data/gte_data_set_with_covariates_and_IPW.RData")
output_dir <- paste0(base_path, "Results")
```

# Loading the data and weight the data with IPW (trim the weights at 30)
We need to define the composite genetic testing awareness variable (we did not keep it in the data frame generated in earlier reports)
```{r}
data_gte <- readRDS(data_file)
nrow(data_gte)

data_gte$aware <- NA
data_gte$aware[which(data_gte$GTE1 == 1 | 
                         data_gte$GTE2 == 1 | 
                         data_gte$GTE3 == 1 | 
                         data_gte$GTE4 == 1)] <- 1
data_gte$aware[which(is.na(data_gte$aware) & 
                         (data_gte$GTE1 == 0 | 
                         data_gte$GTE2 == 0 | 
                         data_gte$GTE3 == 0 | 
                         data_gte$GTE4 == 0))] <- 0

table(data_gte$aware)
survey_gte <- svydesign(id=~PSU_ID, strata=~STRAT, weights=~IPW , data=data_gte)
survey_trim <- trimWeights(survey_gte,upper = 30)
```

# Perform logistic regression of composite measure of genetic testing awareness


```{r}
# we just use the doctor visit variable from visit 1, and not the personal doctor from visit 2. 
model_aware <- svyglm(aware~AGE+SEX+Education+CENTER+Income_level+Current_Health_insurance+Physician_Visit+Background+Language_pref+Marital_status+Employment_status+US_BORN,design=survey_trim,na.action=na.omit,family=quasibinomial())
summary(model_aware)
```

Summarize in a nice table
```{r}
tbl_aware <- tbl_regression(model_aware, exponentiate = TRUE)
tbl_aware %>%
  as_flex_table() %>%
  flextable::save_as_docx(path = paste0(output_dir, "/composite_awarenes_glm.docx"))

```

# Perform separate analysis of each of the awareness questions
We need a supplemental table reporting results for each awareness question separately (rather than the composite awareness variable)
```{r}
# Look at the individual responses:
table(data_gte$GTE1)
table(data_gte$GTE2)
table(data_gte$GTE3)
table(data_gte$GTE4)

# change some "refuse to respond" (values of 9) to NAs. Note that as people respond to more questions, more people refuse to respond. 
# then we need to recreate the survey object.
data_gte$GTE1[which(data_gte$GTE1 == 9)] <- NA
data_gte$GTE2[which(data_gte$GTE2 == 9)] <- NA
data_gte$GTE3[which(data_gte$GTE3 == 9)] <- NA
data_gte$GTE4[which(data_gte$GTE4 == 9)] <- NA

survey_gte <- svydesign(id=~PSU_ID, strata=~STRAT, weights=~IPW , data=data_gte)
survey_trim <- trimWeights(survey_gte,upper = 30)

model_gte1 <- svyglm(GTE1~AGE+SEX+Education+CENTER+Income_level+Current_Health_insurance+Physician_Visit+Background+Language_pref+Marital_status+Employment_status+US_BORN,design=survey_trim,na.action=na.omit,family=quasibinomial())

model_gte2 <- svyglm(GTE2~AGE+SEX+Education+CENTER+Income_level+Current_Health_insurance+Physician_Visit+Background+Language_pref+Marital_status+Employment_status+US_BORN,design=survey_trim,na.action=na.omit,family=quasibinomial())

model_gte3 <- svyglm(GTE3~AGE+SEX+Education+CENTER+Income_level+Current_Health_insurance+Physician_Visit+Background+Language_pref+Marital_status+Employment_status+US_BORN,design=survey_trim,na.action=na.omit,family=quasibinomial())

model_gte4 <- svyglm(GTE4~AGE+SEX+Education+CENTER+Income_level+Current_Health_insurance+Physician_Visit+Background+Language_pref+Marital_status+Employment_status+US_BORN,design=survey_trim,na.action=na.omit,family=quasibinomial())

```

## Summarize the regressions in a table
```{r}
tbl_gte1 <- tbl_regression(model_gte1, exponentiate = TRUE)
tbl_gte2 <- tbl_regression(model_gte2, exponentiate = TRUE)
tbl_gte3 <- tbl_regression(model_gte3, exponentiate = TRUE)
tbl_gte4 <- tbl_regression(model_gte4, exponentiate = TRUE)

tbl_merge_awareness <-
  tbl_merge(
    tbls = list(tbl_gte1, tbl_gte2, tbl_gte3, tbl_gte4),
    tab_spanner = c("**Disease risk**", 
                    "**Risk to children**", 
                    "**Personalized treatment**",
                    "**Drug efficacy**")
  )

ftbl_merge_awarenss <- as_flex_table(tbl_merge_awareness)

# function I found online here
# https://stackoverflow.com/questions/57175351/flextable-autofit-in-a-rmarkdown-to-word-doc-causes-table-to-go-outside-page-mar
FitFlextableToPage <- function(ft, pgwidth = 7){
  ft_out <- ft %>% autofit()
  ft_out <- width(ft_out, width = dim(ft_out)$widths*pgwidth /(flextable_dim(ft_out)$widths))
  return(ft_out)
}

ftbl_merge_awarenss_w <- FitFlextableToPage(ftbl_merge_awarenss)
flextable::save_as_docx(ftbl_merge_awarenss_w, path = paste0(output_dir, "/separate_awarenes_glm.docx"))


```


# Perform another analysis stratified by center. 
We want to see if significant associations in the pooled analysis of all individuals are driven by individuals from a specific site. 

```{r}
subset_site <- subset(survey_trim, CENTER == "Miami")
model_miami <- svyglm(aware~AGE+SEX+Education+Income_level+Current_Health_insurance+Physician_Visit+Background+Language_pref+Marital_status+Employment_status+US_BORN,design=subset_site,na.action=na.omit,family=quasibinomial())

subset_site <- subset(survey_trim, CENTER == "Bronx")
model_bronx <- svyglm(aware~AGE+SEX+Education+Income_level+Current_Health_insurance+Physician_Visit+Background+Language_pref+Marital_status+Employment_status+US_BORN,design=subset_site,na.action=na.omit,family=quasibinomial())

subset_site <- subset(survey_trim, CENTER == "San Diego")
model_sandiego <- svyglm(aware~AGE+SEX+Education+Income_level+Current_Health_insurance+Physician_Visit+Background+Language_pref+Marital_status+Employment_status+US_BORN,design=subset_site,na.action=na.omit,family=quasibinomial())

subset_site <- subset(survey_trim, CENTER == "Chicago")
model_chicago <- svyglm(aware~AGE+SEX+Education+Income_level+Current_Health_insurance+Physician_Visit+Background+Language_pref+Marital_status+Employment_status+US_BORN,design=subset_site,na.action=na.omit,family=quasibinomial())



```

## Summarize the center-specific regressions in a table
```{r}
tbl_miami <- tbl_regression(model_miami, exponentiate = TRUE)
tbl_chicago <- tbl_regression(model_chicago, exponentiate = TRUE)
tbl_bronx <- tbl_regression(model_bronx, exponentiate = TRUE)
tbl_sandiego <- tbl_regression(model_sandiego, exponentiate = TRUE)

tbl_merge_center <-
  tbl_merge(
    tbls = list(tbl_bronx, tbl_chicago, tbl_miami, tbl_sandiego),
    tab_spanner = c("**Bronx**", 
                    "**Chicago**", 
                    "**Miami**",
                    "**San Diego**")
  )

ftbl_merge_center <- as_flex_table(tbl_merge_center)


ftbl_merge_center_w <- FitFlextableToPage(ftbl_merge_center)
flextable::save_as_docx(ftbl_merge_center_w, path = paste0(output_dir, "/awarenes_glm_by_center.docx"))


```

# Because San Diego have mostly Mexicans, redo San Diego analysis without background

```{r}

subset_site <- subset(survey_trim, CENTER == "San Diego")
model_sandiego <- svyglm(aware~AGE+SEX+Education+Income_level+Current_Health_insurance+Physician_Visit+Language_pref+Marital_status+Employment_status+US_BORN,design=subset_site,na.action=na.omit,family=quasibinomial())

tbl_sandiego <- tbl_regression(model_sandiego, exponentiate = TRUE)

tbl_sandiego %>%
  as_flex_table() %>%
  flextable::save_as_docx(path = paste0(output_dir, "/awarenes_glm_by_san_diego_no_background.docx"))


```


# Perform another analysis stratified by gender. 
Another secondary analysis to allow for checking whether some results differ between women and men. 

```{r}
subset_sex <- subset(survey_trim, SEX == "Female")
model_female <- svyglm(aware~AGE+Education+CENTER+Income_level+Current_Health_insurance+Physician_Visit+Background+Language_pref+Marital_status+Employment_status+US_BORN,design=subset_sex,na.action=na.omit,family=quasibinomial())

subset_sex <- subset(survey_trim, SEX == "Male")
model_male <- svyglm(aware~AGE+Education+CENTER+Income_level+Current_Health_insurance+Physician_Visit+Background+Language_pref+Marital_status+Employment_status+US_BORN,design=subset_sex,na.action=na.omit,family=quasibinomial())


tbl_female <- tbl_regression(model_female, exponentiate = TRUE)
tbl_male <- tbl_regression(model_male, exponentiate = TRUE)

tbl_merge_sex <-
  tbl_merge(
    tbls = list(tbl_female, tbl_male),
    tab_spanner = c("**Women**", 
                    "**Men**")
  )

ftbl_merge_sex <- as_flex_table(tbl_merge_sex)


ftbl_merge_sex_w <- FitFlextableToPage(ftbl_merge_sex)
flextable::save_as_docx(ftbl_merge_sex_w, path = paste0(output_dir, "/awarenes_glm_by_gender.docx"))

```



# version:
```{r}
sessionInfo()
```
