---
title: "Computing proportions of individuals aware of genetic tests"
output: 
  html_document:
    toc: true
    keep_md: true
---


```{r}
library(tidyverse)
library(survey)
library(plyr)
library(dplyr)
library(factoextra)
library(labelled)
library(memisc)
library(jtools)
library(Publish)
library(gtsummary)
library(flextable)
library(ggpubr)
```

# Setting up working directory
```{r}
base_path <- "~/Dropbox (Partners HealthCare)/SOL_misc_genetics/survey_gen_test_utilization/20221003_data_code/"
data_file <- paste0(base_path, "Processed_data/gte_data_set_with_covariates_and_IPW.RData")
output_dir <- paste0(base_path, "Results")
```

# Loading the data and weight the data with IPW (trim the weights at 30)
We need to define the composite genetic testing awareness variable (we did not keep it in the data frame generated in earlier reports)
```{r}
data_gte <- readRDS(data_file)
nrow(data_gte)

data_gte$aware <- NA
data_gte$aware[which(data_gte$GTE1 == 1 | 
                         data_gte$GTE2 == 1 | 
                         data_gte$GTE3 == 1 | 
                         data_gte$GTE4 == 1)] <- 1
data_gte$aware[which(is.na(data_gte$aware) & 
                         (data_gte$GTE1 == 0 | 
                         data_gte$GTE2 == 0 | 
                         data_gte$GTE3 == 0 | 
                         data_gte$GTE4 == 0))] <- 0

table(data_gte$aware)
survey_gte <- svydesign(id=~PSU_ID, strata=~STRAT, weights=~IPW , data=data_gte)
survey_trim <- trimWeights(survey_gte,upper = 30)
```

# Compute survey-weighted proportions of awareness of the composite outcome across strata
One of the reviewers has a nice suggestion to make a figure displaying proportions of individuals aware of genetic tests, stratified by education, income, and study site (center).
We here generate the data for that (also generated as part of script 5a) and make a figure (though we will re-create a figure for the paper to match the style of the other paper figure).


```{r}
vars <- c("Education", 
          "CENTER", 
          "Income_level")

res_list <- vector(mode = "list", length = length(vars))

for (i in 1:length(vars)){
  var <- vars[i]
  cur_prop <- svyby(~aware, as.formula(paste0("~", var)),survey_trim, svyciprop)
  cur_est <- round(cur_prop$aware,2)
  cur_se <- cur_prop[["se.as.numeric(aware)"]]
  cur_res <- data.frame(var = var, 
                        stratum = cur_prop[[var]], 
                        prop_aware = cur_est,
                        SE = cur_se,
                        CI_low = cur_est-1.96*cur_se,
                        CI_high = cur_est + 1.96*cur_se)
  res_list[[i]] <- cur_res
}

res <- do.call(rbind, res_list)


```

Save in a file:
```{r}
write.csv(res, file = paste0(output_dir, "/proportion_aware_education_income_center_for_figure.csv"))
```

Figure -- by variables
```{r}
p_edu <- ggplot(res[which(res$var == "Education"),], aes(x=stratum, y=prop_aware, fill=stratum)) + 
      geom_bar(stat="identity", position=position_dodge(0.9))+ 
      geom_errorbar(aes(ymin=CI_low, ymax=CI_high), width=.2, position=position_dodge(.9)) +
      scale_fill_manual(values=c("#fec44f", "#2b8cbe","#a6bddb", "#2ca25f")) +
      theme_minimal() + xlab("") + ylab("Proportion aware") + 
      theme(legend.position = "none") +
      theme(axis.text.x = element_text(angle = 60, vjust=1, hjust =1)) + coord_flip()

p_center <- ggplot(res[which(res$var == "CENTER"),], aes(x=stratum, y=prop_aware, fill=stratum)) + 
      geom_bar(stat="identity", position=position_dodge(0.9))+ 
      geom_errorbar(aes(ymin=CI_low, ymax=CI_high), width=.2, position=position_dodge(.9)) +
      scale_fill_manual(values=c("#fec44f", "#2b8cbe","#a6bddb", "#2ca25f")) +
      theme_minimal() + xlab("") + ylab("Proportion aware") + 
      theme(legend.position = "none") +
      theme(axis.text.x = element_text(angle = 60, vjust=1, hjust =1)) + coord_flip()

p_income <- ggplot(res[which(res$var == "Income_level"),], aes(x=stratum, y=prop_aware, fill=stratum)) + 
      geom_bar(stat="identity", position=position_dodge(0.9))+ 
      geom_errorbar(aes(ymin=CI_low, ymax=CI_high), width=.2, position=position_dodge(.9)) +
      scale_fill_manual(values=c("#fec44f", "#2b8cbe","#a6bddb", "#2ca25f", "#99d8c9")) +
      theme_minimal() + xlab("") + ylab("Proportion aware") + 
      theme(legend.position = "none") +
      theme(axis.text.x = element_text(angle = 60, vjust=1, hjust =1)) + coord_flip()


fig <- ggarrange(p_edu, p_center, p_income, ncol = 1, align = "v")

ggexport(fig, filename = file.path(output_dir, "awareness_edu_enter_income.pdf"))

```



# version:
```{r}
sessionInfo()
```
