---
title: "Poisson regression of perceived utility of genetic tests"
output: 
  html_document:
    toc: true
    keep_md: true
---

```{r,message=FALSE}
library(tidyverse)
library(survey)
library(plyr)
library(dplyr)
library(factoextra)
library(labelled)
library(tableone)
library(memisc)
library(jtools)
library(Publish)
library(gtsummary)
library(flextable)
```



# Setting up working directory
```{r}
base_path <- "~/Dropbox (Partners HealthCare)/SOL_misc_genetics/survey_gen_test_utilization/20210917_data_code/"
data_file <- paste0(base_path, "Processed_data/20210918_gte_data_set_with_covariates_and_IPW.RData")
output_dir <- paste0(base_path, "Results")
```

# Load the data and weight it with IPW (trim the weight at 30)

```{r}
data_gte <- readRDS(data_file)
nrow(data_gte)
hist(data_gte$GTE5)

survey_gte <- svydesign(id=~PSU_ID, strata=~STRAT, weights=~IPW , data=data_gte)
survey_trim <- trimWeights(survey_gte,upper = 30)
```
There are `{r sum(is.na(data_gte$GTE5))}` missing values in the response to the perceived utility question. 


# Perform possion regression of percieved utility.
The responses to the perceived utility question are roughly log-normal. We use Poisson regression, modeling multiplicative effects of the covariates. 

```{r}
model_utility <- svyglm(GTE5~AGE+SEX+Education+CENTER+Income_level+Current_Health_insurance+Physician_Visit+Background+Language_pref+Marital_status+Employment_status+US_BORN,design=survey_trim,na.action=na.omit,family=quasipoisson())
summary(model_utility)
```

Summarize in a nice table
```{r}
tbl_utility <- tbl_regression(model_utility, exponentiate = TRUE)
tbl_utility %>%
  as_flex_table() %>%
  flextable::save_as_docx(path = paste0(output_dir, "/20210918_utility_poisson_reg.docx"))

```

# Secondary analysis: stratify by center
This analysis will allow others, if interested, to check whether some results seem to be driven by a specific center (though not to test this hypothesis).

```{r}
subset_site <- subset(survey_trim, CENTER == "Miami")
model_miami <- svyglm(GTE5~AGE+SEX+Education+Income_level+Current_Health_insurance+Physician_Visit+Background+Language_pref+Marital_status+Employment_status+US_BORN,design=subset_site,na.action=na.omit,family=quasipoisson())

subset_site <- subset(survey_trim, CENTER == "Bronx")
model_bronx <- svyglm(GTE5~AGE+SEX+Education+Income_level+Current_Health_insurance+Physician_Visit+Background+Language_pref+Marital_status+Employment_status+US_BORN,design=subset_site,na.action=na.omit,family=quasipoisson())

subset_site <- subset(survey_trim, CENTER == "San Diego")
model_sandiego <- svyglm(GTE5~AGE+SEX+Education+Income_level+Current_Health_insurance+Physician_Visit+Background+Language_pref+Marital_status+Employment_status+US_BORN,design=subset_site,na.action=na.omit,family=quasipoisson())

subset_site <- subset(survey_trim, CENTER == "Chicago")
model_chicago <- svyglm(GTE5~AGE+SEX+Education+Income_level+Current_Health_insurance+Physician_Visit+Background+Language_pref+Marital_status+Employment_status+US_BORN,design=subset_site,na.action=na.omit,family=quasipoisson())

# merge results and prepare a combined table:

tbl_miami <- tbl_regression(model_miami, exponentiate = TRUE)
tbl_chicago <- tbl_regression(model_chicago, exponentiate = TRUE)
tbl_bronx <- tbl_regression(model_bronx, exponentiate = TRUE)
tbl_sandiego <- tbl_regression(model_sandiego, exponentiate = TRUE)

tbl_merge_center <-
  tbl_merge(
    tbls = list(tbl_bronx, tbl_chicago, tbl_miami, tbl_sandiego),
    tab_spanner = c("**Bronx**", 
                    "**Chicago**", 
                    "**Miami**",
                    "**San Diego**")
  )

ftbl_merge_center <- as_flex_table(tbl_merge_center)


ftbl_merge_center_w <- FitFlextableToPage(ftbl_merge_center)
flextable::save_as_docx(ftbl_merge_center_w, path = paste0(output_dir, "/20210918_perceived utility_reg_by_center.docx"))



```


# Secondary analysis: stratify by center
This analysis will allow others, if interested, to check whether some results seem to be driven by a specific center (though not to test this hypothesis).

```{r}
subset_sex <- subset(survey_trim, SEX == "Female")
model_female <- svyglm(GTE5~AGE+Education+CENTER+Income_level+Current_Health_insurance+Physician_Visit+Background+Language_pref+Marital_status+Employment_status+US_BORN,design=subset_sex,na.action=na.omit,family=quasipoisson())

subset_sex <- subset(survey_trim, SEX == "Male")
model_male <- svyglm(GTE5~AGE+Education+CENTER+Income_level+Current_Health_insurance+Physician_Visit+Background+Language_pref+Marital_status+Employment_status+US_BORN,design=subset_sex,na.action=na.omit,family=quasipoisson())


tbl_female <- tbl_regression(model_female, exponentiate = TRUE)
tbl_male <- tbl_regression(model_male, exponentiate = TRUE)

tbl_merge_sex <-
  tbl_merge(
    tbls = list(tbl_female, tbl_male),
    tab_spanner = c("**Women**", 
                    "**Men**")
  )

ftbl_merge_sex <- as_flex_table(tbl_merge_sex)


ftbl_merge_sex_w <- FitFlextableToPage(ftbl_merge_sex)
flextable::save_as_docx(ftbl_merge_sex_w, path = paste0(output_dir, "/20210918_perceived_utility_reg_by_gender.docx"))

```

```{r}
sessionInfo()
```

