---
title: "Usings contrasts to provide effect estimates and CIs effects we may want to report"
output:
  html_document:
    toc: yes
    keep_md: true
  pdf_document:
    toc: yes
---

```{r,message=FALSE}
library(tidyverse)
library(survey)
library(plyr)
library(dplyr)
library(factoextra)
library(labelled)
library(tableone)
library(memisc)
library(jtools)
library(Publish)
library(dummies)
```
# Setting up working directory
```{r}
base_path <- "~/Dropbox (Partners HealthCare)/SOL_misc_genetics/survey_gen_test_utilization/20210917_data_code/"
data_file <- paste0(base_path, "Processed_data/20210918_gte_data_set_with_covariates_and_IPW.RData")
```


# Computing effect estimates for significant differences
For every effect estimate that had p-value<0.05, we here compute an estimate (making sure we are comparing "extreme" factors) and 95% confidence interval. We also compute (weighted) proportions. This is useful because we can use this to put results in the text. 

# Load the data and prepare the model

```{r}
data_gte <- readRDS(data_file)
nrow(data_gte)

data_gte$aware <- NA
data_gte$aware[which(data_gte$GTE1 == 1 | 
                         data_gte$GTE2 == 1 | 
                         data_gte$GTE3 == 1 | 
                         data_gte$GTE4 == 1)] <- 1
data_gte$aware[which(is.na(data_gte$aware) & 
                         (data_gte$GTE1 == 0 | 
                         data_gte$GTE2 == 0 | 
                         data_gte$GTE3 == 0 | 
                         data_gte$GTE4 == 0))] <- 0

table(data_gte$aware)

data_gte$GTE1[which(data_gte$GTE1 == 9)] <- NA
data_gte$GTE2[which(data_gte$GTE2 == 9)] <- NA
data_gte$GTE3[which(data_gte$GTE3 == 9)] <- NA
data_gte$GTE4[which(data_gte$GTE4 == 9)] <- NA

```



# Prreparing dummy variables of interest
Creating dummy variables in preparation for computing contrast
```{r}
data_gte$income_10000 <- ifelse(data_gte$Income_level=="Less than $10,000",1,0)
data_gte$income_20000 <- ifelse(data_gte$Income_level=="$10,001-$20,000",1,0)
data_gte$income_40000 <- ifelse(data_gte$Income_level=="$20,001-$40,000",1,0)
data_gte$income_75000 <- ifelse(data_gte$Income_level=="$40,001-$75,000",1,0)
data_gte$income_75000_high <- ifelse(data_gte$Income_level=="More than $75,000",1,0)

data_gte$Pref_english <- ifelse(data_gte$Language_pref == "English", 1, 0)
data_gte$Pref_spanish <- ifelse(data_gte$Language_pref == "Spanish", 1, 0)

data_gte$age_40 <- ifelse(data_gte$AGE=="<40",1,0)
data_gte$age_60 <- ifelse(data_gte$AGE=="41-60",1,0)
data_gte$age_60_high <- ifelse(data_gte$AGE=="61<",1,0)

data_gte$us_born_yes <- ifelse(data_gte$US_BORN=="Yes",1,0)
data_gte$us_born_no <- ifelse(data_gte$US_BORN=="No",1,0)

data_gte$center_bronx <- ifelse(data_gte$CENTER=="Bronx",1,0)
data_gte$center_miami <- ifelse(data_gte$CENTER=="Miami",1,0)
data_gte$center_chicago <- ifelse(data_gte$CENTER=="Chicago",1,0)
data_gte$center_sandiego <- ifelse(data_gte$CENTER=="San Diego",1,0)

# marital status
data_gte$Marital_status_single <- ifelse(data_gte$Marital_status=="Single",1,0)
data_gte$Marital_status_partner <- ifelse(data_gte$Marital_status=="Married or living with a partner",1,0)
data_gte$Marital_status_separate <- ifelse(data_gte$Marital_status=="Separated,divorced,or widow(er)",1,0)



# education level
data_gte$edu_less_12 <- ifelse(data_gte$Education=="<12",1,0)
data_gte$edu_12 <- ifelse(data_gte$Education=="12",1,0)
data_gte$edu_more_12 <- ifelse(data_gte$Education==">12",1,0)

survey_gte <- svydesign(id=~PSU_ID, strata=~STRAT, weights=~IPW , data=data_gte)
survey_trim <- trimWeights(survey_gte,upper = 30)
```

# Create a function of computing p_value from chi-square test
```{r}
contrast_OR_pval <- function(contrast, model){
  
  contrast <- as.data.frame(contrast)[1,]
	t_stat <- as.numeric(abs(contrast[1]/contrast[2]))
	OR <- as.numeric(exp(contrast[1]))
	df <- model$df.residual
	Z <- qt(0.975,df = df, ncp=0, lower.tail = TRUE)
	pval <- (1-pt(t_stat, df =df, lower.tail = TRUE))*2
	CI <- c(as.numeric(contrast[1] - Z*contrast[2]), as.numeric(contrast[1] + Z*contrast[2]))
	return(list(OR= round(OR,2), pval = pval, CI=  round(exp(CI),2)))
}
```

## Education comparison
```{r}
model_aware_edu <- svyglm(aware~AGE+SEX+edu_less_12+edu_more_12 + 
                                Income_level+ 
                                CENTER  + 
                               Current_Health_insurance+
                               Physician_Visit+Background+Language_pref+
                               Marital_status+Employment_status+US_BORN,
                             design=survey_trim,na.action=na.omit,family=quasibinomial())
contrast <- svycontrast(model_aware_edu, list(diff=c(edu_more_12=1, edu_less_12=-1)))
contrast_OR_pval(contrast,model_aware_edu)

# also compute proportion of awareness by education groups:
svyby(~aware, ~Education,survey_trim, svyciprop)

```


## Miami and Chicago comparison
```{r}
model_aware_center <- svyglm(aware~AGE+SEX+Education+
                                Income_level+ 
                                center_chicago +
                               center_sandiego + center_miami + 
                               Current_Health_insurance+
                               Physician_Visit+Background+Language_pref+
                               Marital_status+Employment_status+US_BORN,
                             design=survey_trim,na.action=na.omit,family=quasibinomial())
contrast <- svycontrast(model_aware_center, list(diff=c(center_miami=1, center_chicago=-1)))
contrast_OR_pval(contrast,model_aware_center)

# just reverse the order...
contrast <- svycontrast(model_aware_center, list(diff=c(center_chicago=-1, center_miami=1)))
contrast_OR_pval(contrast,model_aware_center)

# compute proportions:
svyby(~aware, ~CENTER,survey_trim, svyciprop)
```

Number of complete cases used in the analysis: `r length(model_aware_center$y)`.

## Marital status
```{r}
model_drug_marital <- svyglm(aware~AGE+SEX+Education+
                                Income_level+ 
                                CENTER + Language_pref  + 
                               Marital_status_separate + Marital_status_single + 
                               Current_Health_insurance+
                               Physician_Visit+Background+
                               Marital_status+Employment_status+US_BORN,
                            design=survey_trim,na.action=na.omit,family=quasibinomial())

contrast <- svycontrast(model_drug_marital, list(diff=c(Marital_status_separate=1, Marital_status_single=-1)))
contrast_OR_pval(contrast,model_drug_marital)

# compute proportions:
svyby(~aware, ~Marital_status,survey_trim, svyciprop)
```

## Income comparison
```{r}
model_aware_income <- svyglm(aware~AGE+SEX+Education+
                                income_10000 + income_75000_high + 
                               income_75000 + income_20000 + 
                                CENTER + 
                               Current_Health_insurance+
                               Physician_Visit+Background+Language_pref+
                               Marital_status+Employment_status+US_BORN,
                             design=survey_trim,na.action=na.omit,family=quasibinomial())
contrast <- svycontrast(model_aware_income, list(diff=c(income_75000_high=1, income_10000=-1)))
contrast_OR_pval(contrast,model_aware_income)

# compute proportions:
svyby(~aware, ~Income_level,survey_trim, svyciprop)

```


# Contrasts based on specific genetic tests


## Drug efficacy: separated, divorced or widdower, versus married or living with a partner.

```{r}
model_drug_marital <- svyglm(GTE4~AGE+SEX+Education+
                                Income_level+ 
                                CENTER + Language_pref  + 
                               Marital_status_separate + Marital_status_partner + 
                               Current_Health_insurance+
                               Physician_Visit+Background+
                               Marital_status+Employment_status+US_BORN,
                            design=survey_trim,na.action=na.omit,family=quasibinomial())

contrast <- svycontrast(model_drug_marital, list(diff=c(Marital_status_separate=1, Marital_status_partner=-1)))
contrast_OR_pval(contrast,model_drug_marital)

# compute proportions:
svyby(~GTE4, ~Marital_status,survey_trim, svyciprop)
```

# separate versus single

```{r}
model_drug_marital <- svyglm(GTE4~AGE+SEX+Education+
                                Income_level+ 
                                CENTER + Language_pref  + 
                               Marital_status_separate + Marital_status_single + 
                               Current_Health_insurance+
                               Physician_Visit+Background+
                               Employment_status+US_BORN,
                            design=survey_trim,na.action=na.omit,family=quasibinomial())

contrast <- svycontrast(model_drug_marital, list(diff=c(Marital_status_separate=1, Marital_status_single=-1)))
contrast_OR_pval(contrast,model_drug_marital)

# compute proportions:
svyby(~GTE4, ~Marital_status,survey_trim, svyciprop)
```
## Compute proportions of population groups with awareness - disease risk

```{r}

# by sex
svyby(~GTE1, ~SEX,survey_trim, svyciprop)

# by education
svyby(~GTE1, ~Education,survey_trim, svyciprop)

# by center
svyby(~GTE1, ~CENTER,survey_trim, svyciprop)


# by income
svyby(~GTE1, ~Income_level,survey_trim, svyciprop)

# by background
svyby(~GTE1, ~Background,survey_trim, svyciprop)
```
## Compute proportions of population groups with awareness - risk to children

```{r}

# by sex
svyby(~GTE2, ~SEX,survey_trim, svyciprop)

# by education
svyby(~GTE2, ~Education,survey_trim, svyciprop)

# by center
svyby(~GTE2, ~CENTER,survey_trim, svyciprop)


# by income
svyby(~GTE2, ~Income_level,survey_trim, svyciprop)

# by background
svyby(~GTE2, ~Background,survey_trim, svyciprop)
```



## Compute proportions of population groups with awareness - personalized treatment

```{r}

# by sex
svyby(~GTE3, ~SEX,survey_trim, svyciprop)

# by education
svyby(~GTE3, ~Education,survey_trim, svyciprop)

# by center
svyby(~GTE3, ~CENTER,survey_trim, svyciprop)


# by income
svyby(~GTE3, ~Income_level,survey_trim, svyciprop)

# by background
svyby(~GTE3, ~Background,survey_trim, svyciprop)

# by age
svyby(~GTE3, ~AGE,survey_trim, svyciprop)


# by employment status
svyby(~GTE3, ~Employment_status,survey_trim, svyciprop)
```




## Compute proportions of population groups with awareness - drug efficacy

```{r}

# by sex
svyby(~GTE4, ~SEX,survey_trim, svyciprop)

# by education
svyby(~GTE4, ~Education,survey_trim, svyciprop)

# by center
svyby(~GTE4, ~CENTER,survey_trim, svyciprop)


# by income
svyby(~GTE4, ~Income_level,survey_trim, svyciprop)

# by background
svyby(~GTE4, ~Background,survey_trim, svyciprop)

# by age
svyby(~GTE4, ~AGE,survey_trim, svyciprop)


# by employment status
svyby(~GTE4, ~Employment_status,survey_trim, svyciprop)

# by language preference
svyby(~GTE4, ~Language_pref,survey_trim, svyciprop)
```



